---
title: "HW1 Part 1"
author: "Sibei Zhang, Richard Tang & Alexander Quispe"
date: "`r Sys.Date()`"
output: html_document
---
```{r lib, include=FALSE}
library(dplyr)
library(haven)
library(tidyverse)
library(lmtest)
library(sandwich)
library(grf)
library(glmnet)
library(splines)
library(ggplot2)
library(reshape2)
```


```{r}
star <- read_sav("C:/Users/Alexander/Documents/GitHub/ECO293/data/Project_STAR/PROJECT_STAR/STAR_Students.sav")
names(star)<-tolower(names(star))
options(stringAsFactors=FALSE)
```

```{r drop_relabel}
star <- star[!(star$g1classtype== 3 ),]
star$g1classtype[star$g1classtype == 1] <- 1
star$g1classtype[star$g1classtype == 2] <- 0
```

# 1. average treatment effect in the randomized experiment

```{r ate_ci}
#ate
ate_est <- mean(star$g1tmathss[star$g1classtype==1], na.rm = TRUE) - mean(star$g1tmathss[star$g1classtype==0], na.rm = TRUE)

ate_est

#confidence intervals
ate_se <- sqrt(var(star$g1tmathss[star$g1classtype == 1], na.rm = TRUE) / sum(star$g1classtype == 1, na.rm = TRUE) + var(star$g1tmathss[star$g1classtype == 0], na.rm = TRUE) / sum(star$g1classtype == 0, na.rm = TRUE))

ate_se

print(paste0("95% CI: ", round(ate_est),
" +/- ", round(1.96 * ate_se)))
```


# 2 Describe your method(s) for systematically deleting some observations as a function of X’s

```{r drop_obs}
names(star)[names(star) == 'g1classtype'] <- "W"
# copying old dataset, just in case
data.exp <- star
data <- star

# defining the group that we will be dropped with some high probability
grp <- ((data$W == 1) &  # if treated AND...
        (
            (data$race == 2) |     # belongs an older group OR
            (data$g4surban == 3)   # more conservative
        )) | # OR
        ((data$W == 0) &  # if untreated AND...
        (
            (data$race == 1 ) |     # belongs a younger group OR
            (data$g4surban == 4)   # more liberal
        )) 

# Individuals in the group above have a small chance of being kept in the sample
prob.keep <- ifelse(grp, .15, .85)
keep.idx <- as.logical(rbinom(n=nrow(data), prob=prob.keep, size = 1))
# Dropping
star <- data[keep.idx,]
```

# 2.1 new difference-in-means point estimate of the treatment effect is significantly different

```{r new_ate}
#ate
ate_est <- mean(star$g1tmathss[star$W==1], na.rm = TRUE) - mean(star$g1tmathss[star$W==0], na.rm = TRUE)

ate_est

#confidence intervals
ate_se <- sqrt(var(star$g1tmathss[star$W == 1], na.rm = TRUE) / sum(star$W == 1, na.rm = TRUE) + var(star$g1tmathss[star$W == 0], na.rm = TRUE) / sum(star$W == 0, na.rm = TRUE))

ate_se

print(paste0("95% CI: ", round(ate_est),
" +/- ", round(1.96 * ate_se)))
```

# 2.2 Propensity Score


```{r pscore}
star_2 <- star[complete.cases(star$gender, star$race, star$flagsgk, star$gktrace, star$g4surban, star$W, star$g1tmathss),]


# Additional covariates
covariates <- c("gender", "race", "flagsgk", "gktrace", "g4surban")

# Estimate the propensity score e(X) via logistic regression using splines
fmla <- as.formula(paste0("~", paste0("bs(", covariates, ", df=3)", collapse="+")))
Z <- star_2$W
Y <- star_2$g1tmathss
XX <- model.matrix(fmla, star_2)
logit <- cv.glmnet(x=XX, y=Z, family="binomial")
e.hat <- predict(logit, XX, s = "lambda.min", type="response")

# Histogram 
hist(e.hat) 

```

# 3.1 The difference in means estimator (which would be appropriate in an RCT).

```{r ate_pscore}
#ate
ate_est <- mean(star_2$g1tmathss[star_2$W==1], na.rm = TRUE) - mean(star_2$g1tmathss[star_2$W==0], na.rm = TRUE)

ate_est

#confidence intervals
ate_se <- sqrt(var(star_2$g1tmathss[star_2$W == 1], na.rm = TRUE) / sum(star_2$W == 1, na.rm = TRUE) + var(star_2$g1tmathss[star_2$W == 0], na.rm = TRUE) / sum(star_2$W == 0, na.rm = TRUE))

ate_se

print(paste0("95% CI: ", round(ate_est),
" +/- ", round(1.96 * ate_se)))

```

# 3.2  Simple linear regression, i.e., fit Y ~ X + W and interpret the W-coefficient as the ATE.


```{r ols}
fmla <- g1tmathss ~ (W + gender + race + flagsgk + gktrace+ g4surban)
regbasic <- lm(fmla, data=star_2)
summary(regbasic) 
ols <- lm(fmla, data=star_2)
coeftest(ols, vcov=vcovHC(ols, type='HC2'))[2,]
```


# 3.3 AIPW with cross-fitting, using a machine learning method of your choice.

```{r aipw}
data <- star_2
# Available in randomized settings and observational settings with unconfoundedness+overlap

# A list of vectors indicating the left-out subset
n <- nrow(data)
n.folds <- 5
indices <- split(seq(n), sort(seq(n) %% n.folds))

# Preparing data
W <- data$W
Y <- data$g1tmathss
covariates <- c("gender", "race", "flagsgk", "gktrace", "g4surban")

# Matrix of (transformed) covariates used to estimate E[Y|X,W]
fmla.xw <- formula(paste("~ 0 +", paste0("bs(", covariates, ", df=3)", "*", W, collapse=" + ")))
XW <- model.matrix(fmla.xw, data)
# Matrix of (transformed) covariates used to predict E[Y|X,W=w] for each w in {0, 1}
data.1 <- data
data.1$W <- 1
XW1 <- model.matrix(fmla.xw, data.1)  # setting W=1
data.0 <- data
data.0$W <- 0
XW0 <- model.matrix(fmla.xw, data.0)  # setting W=0

# Matrix of (transformed) covariates used to estimate and predict e(X) = P[W=1|X]
fmla.x <- formula(paste(" ~ 0 + ", paste0("bs(", covariates, ", df=3)", collapse=" + ")))
XX <- model.matrix(fmla.x, data)


# # (Optional) Not penalizing the main effect (the coefficient on W)
penalty.factor <- rep(1, ncol(XW))
 #penalty.factor[colnames(XW) == W] <- 0

# Cross-fitted estimates of E[Y|X,W=1], E[Y|X,W=0] and e(X) = P[W=1|X]
mu.hat.1 <- rep(NA, n)
mu.hat.0 <- rep(NA, n)
e.hat <- rep(NA, n)
for (idx in indices) {
  # Estimate outcome model and propensity models
  # Note how cross-validation is done (via cv.glmnet) within cross-fitting! 
  outcome.model <- cv.glmnet(x=XW[-idx,], y=Y[-idx], family="gaussian", penalty.factor=penalty.factor)
  propensity.model <- cv.glmnet(x=XX[-idx,], y=W[-idx], family="binomial")

  # Predict with cross-fitting
  mu.hat.1[idx] <- predict(outcome.model, newx=XW1[idx,], type="response")
  mu.hat.0[idx] <- predict(outcome.model, newx=XW0[idx,], type="response")
  e.hat[idx] <- predict(propensity.model, newx=XX[idx,], type="response")
}

# Commpute the summand in AIPW estimator
aipw.scores <- (mu.hat.1 - mu.hat.0
                + W / e.hat * (Y -  mu.hat.1)
                - (1 - W) / (1 - e.hat) * (Y -  mu.hat.0))

# Tally up results
ate.aipw.est <- mean(aipw.scores)
ate.aipw.se <- sd(aipw.scores) / sqrt(n)
ate.aipw.tstat <- ate.aipw.est / ate.aipw.se
ate.aipw.pvalue <- 2*(pnorm(1 - abs(ate.aipw.tstat)))
ate.aipw.results <- c(estimate=ate.aipw.est, std.error=ate.aipw.se, t.stat=ate.aipw.tstat, pvalue=ate.aipw.pvalue)
print(ate.aipw.results)
```
