---
title: "HW1 Part II-1"
author: "Sibei Zhang, Richard Tang & Alexander Quispe"
date: "`r Sys.Date()`"
output: html_document
---
```{r lib, include=FALSE}
library(dplyr)
library(haven)
library(tidyverse)
library(lmtest)
library(sandwich)
library(grf)
library(glmnet)
library(splines)
library(ggplot2)
library(reshape2)
```


```{r}
star <- read_sav("C:/Users/Alexander/Documents/GitHub/ECO293/data/Project_STAR/PROJECT_STAR/STAR_Students.sav")
names(star)<-tolower(names(star))
options(stringAsFactors=FALSE)
```

```{r drop_relabel}
star <- star[!(star$g1classtype== 3 ),]
star$g1classtype[star$g1classtype == 1] <- 1
star$g1classtype[star$g1classtype == 2] <- 0
```

# 1. average treatment effect in the randomized experiment

```{r ate_ci}
#ate
ate_est <- mean(star$g1tmathss[star$g1classtype==1], na.rm = TRUE) - mean(star$g1tmathss[star$g1classtype==0], na.rm = TRUE)

ate_est

#confidence intervals
ate_se <- sqrt(var(star$g1tmathss[star$g1classtype == 1], na.rm = TRUE) / sum(star$g1classtype == 1, na.rm = TRUE) + var(star$g1tmathss[star$g1classtype == 0], na.rm = TRUE) / sum(star$g1classtype == 0, na.rm = TRUE))

ate_se

print(paste0("95% CI: ", round(ate_est),
" +/- ", round(1.96 * ate_se)))
```


# 2 Describe your method(s) for systematically deleting some observations as a function of X’s

```{r drop_obs}
names(star)[names(star) == 'g1classtype'] <- "W"
# copying old dataset, just in case
data.exp <- star
data <- star

# defining the group that we will be dropped with some high probability
grp <- ((data$W == 1) &  # if treated AND...
        (
            (data$race == 2) |     # belongs an older group OR
            (data$g4surban == 3)   # more conservative
        )) | # OR
        ((data$W == 0) &  # if untreated AND...
        (
            (data$race == 1 ) |     # belongs a younger group OR
            (data$g4surban == 4)   # more liberal
        )) 

# Individuals in the group above have a small chance of being kept in the sample
prob.keep <- ifelse(grp, .15, .85)
keep.idx <- as.logical(rbinom(n=nrow(data), prob=prob.keep, size = 1))
# Dropping
star <- data[keep.idx,]
```

# Part II: Heterogeneous Treatment Effects in Observational Studies.
We start by considering different random forest based strategies for estimation heterogeneous
treatment effects (HTE).

## Use the regression_forest function to estimate HTEs via the S-learner strategy.


```{r reg_forest}
star_2 <- star[complete.cases(star$gender, star$race, star$flagsgk, star$gktrace, star$g4surban, star$W, star$g1tmathss),]


# Additional covariates
covariates <- c("gender", "race", "flagsgk", "gktrace", "g4surban")

# Estimate the propensity score e(X) via logistic regression using splines
fmla <- as.formula(paste0("~", paste0("bs(", covariates, ", df=3)", collapse="+")))
W <- star_2$W
Y <- star_2$g1tmathss

Y <-star_2[['g1tmathss']]
X <- model.matrix(fmla, star_2)
#Z = cbind(X, W)

Y <- 5*rnorm(nrow(X))



sf = regression_forest(cbind(X, W), Y)

```











